---
title: "Life Insurance Simulation"
author: "Andy Merlino"
date: "June 4, 2015"
output: html_document
---


We provide whole life insurance.  

How much money do we need to set aside today to have a probability, p, that we can payout the entire life insurance  benefit to all insurees?

```{r load_stuff, echo = FALSE}
library(lifetable) # ractuary package
library(knitr)

policies <- read.csv("policies.csv")

options(scipen=999)
```

These are the life insurance policies we insure.

```{r view_policies, echo=FALSE, results = "asis"}
kable(policies)
```

As you can see, we have male and female insurees of varying ages.  They also have varying benefits.

We will use the official 2010 US social security life table as provided on the [Official Social Security Website](http://www.ssa.gov/oact/STATS/table4c6.html).

```{r life_tables, echo = FALSE}
rate <- 0.04
n <- 1000

qx_male <- LifeTable(x = qx_data$x, q_x = qx_data$male_qx)
qx_male <- ActuarialTable(i = rep(rate, times = length(qx_male@x)), qx_male)

qx_female <- LifeTable(x = qx_data$x, q_x = qx_data$female_qx)
qx_female <- ActuarialTable(i = rep(rate, times = length(qx_female@x)), qx_female)
```

We simulate each individuals death in accordance with the multinomial distribution specified by either the male or femal qx table and the individual's age.  We run `r n` observations of each individual's future life expectancy.  We assume each individual future life expectancy is independent and identically distributed.  We discount the death benefit, assuming a `r rate * 100` % interest rate, for the number of years the insuree survives.  All death benefits are paid at the end of the year of death.

```{r}
# use lifetable package to define each policy as a Z_x object
hold <- list()
for (j in 1:nrow(policies)) {
  hold[[j]] <- if (policies$gender[j] == "male") {
    lifetable::Z_x(x_ = policies$age[j], 
                    t_ = max(qx_male@x) - policies$age[j],
                    benefit = policies$policy[j], 
                    qx_male)
  } else {
    lifetable::Z_x(x_ = policies$age[j], 
                    t_ = max(qx_female@x) - policies$age[j],
                    benefit = policies$policy[j], 
                    qx_female)
  }
}

# make simulation a lifetable generic\
out <- lapply(hold, function(k) lifetable::rdeath(k, n = n)$pv)
out <- matrix(unlist(out), ncol = n, byrow = TRUE)
total <- apply(out, 2, sum)
c(mean = mean(total), quantile(total, seq(0.70, 0.95, by = 0.05)))
```