---
title: "Life Insurance Simulation"
author: "Andy Merlino"
date: "June 4, 2015"
output: html_document
---


We provide whole like insurance.  How much money do we need to set aside today to have a 90% chance that we can payout the entire benefit at the end of each year that each individual dies.

```{r load_stuff, echo = FALSE}
library(lifetable) # ractuary package
library(knitr)

policies <- read.csv("policies.csv")

options(scipen=999)
```

Here is what the policies we currently insure look like.

```{r view_policies, echo=FALSE, results = "asis"}
kable(policies)
```

As you can see, we have male and female insurees of varying ages.  They also have varying benefits.

We are using the official 2010 US social security life table.  This is the table's structure:

```{r view_qx_table}
str(lifetable::qx_data)
```

Classify each policy as a Z_x object

```{r}
qx_male <- LifeTable(x = qx_data$x, q_x = qx_data$male_qx)
qx_male <- ActuarialTable(i = rep(0.04, times = length(qx_male@x)), qx_male)

qx_female <- LifeTable(x = qx_data$x, q_x = qx_data$female_qx)
qx_female <- ActuarialTable(i = rep(0.04, times = length(qx_female@x)), qx_female)
```

Simulate
```{r}
# find the probability of dieing at each age
#qx, tpx
test <- Z_x(x_ = policies$age[2], payable = policies$policy[2], qx_male)

# make simulation a lifetable generic
mqx <- list(vector, length(test@qx[x_:length()]))
mqx[1] <- test@qx[1]
for (j in 2:length(qx)) {
    mqx[j] <- tpx[j - 1] * qx[j]
}
unlist(mqx)
```