---
title: "ChainLadder Package in Action"
author: "Andy Merlino"
date: "Tuesday, November 18, 2014"
output: html_document
---

### Purpose

The `ChainLadder` package includes methods for projecting insurance loss reserves.  For more information please see the [package vignette](http://cran.r-project.org/web/packages/ChainLadder/vignettes/ChainLadder.pdf).  This document aims to measure the accuracy of the different reserving methods available in the `ChainLadder` package by comparing the package projections to historical data.

### Data

The CAS website provides historic U.S. insurance loss data accross 6 different lines of business and many different companies.  The data includes complete triangles from origin years 1988 through 1997.  We will use data from origin years 1988 through 1997 evaluated as of 1997 to run the `ChainLadder` projections.  The data will be projected out to the 10th development year and compared to actual loss data at that time.  The CAS data sets include data from the following lines of business:


For more information on the data see the [CAS webpage](http://www.casact.org/research/index.cfm?fa=loss_reserves_data).

```{r packages_data, message = FALSE}
# load the ChainLadder package
library(ChainLadder) # install.packages("ChainLadder")
# load the CAS data
library(casdata) # devtools::install_github("merlinoa/casdata")
library(exhibit)
library(pander)
```

### A little data cleaning

A good portion of the companies in the data sets are missing loss information for a portion of the origin years (e.g for workers compensation 70 out of 132).  We are going to remove these companies from our analysis.

```{r data_cleaning}
# Combine data sets into a list 
data_sets <- list(wkcomp, ppauto, prodliab, comauto, medmal, othliab)

# identiy unique companies in each data set that have missing data
# I also had some trouble with companies that had very small losses, so I decided
# to remove all companies with losses less than 100 at any observation from the data set.
id_missing_data <- lapply(data_sets, function(x) unique(x$GRCODE[x[, 6] <= 100]))

# remove companies that have missing data
for (i in seq_along(data_sets)) {
  data_sets[[i]] <- data_sets[[i]][-which(data_sets[[i]]$GRCODE %in% id_missing_data[[i]]), ]
}
```

### Workers' Compensation

First we need to make the triangles.  We will look at the paid triangles first.

```{r wc_paid_tri}
wkcomp_clean <- data_sets[[1]]

wkcomp_id <- unique(wkcomp_clean$GRCODE)

paid_tri <- vector("list", length(wkcomp_id))

for (i in seq_along(wkcomp_id)) {
  paid_tri[[i]] <- as.triangle(wkcomp_clean[wkcomp_clean$GRCODE == wkcomp_id[i] & 
                                      wkcomp_clean$AccidentYear + wkcomp_clean$DevelopmentLag < 1999, ], 
                               origin = "AccidentYear", 
                               dev = "DevelopmentLag", 
                               value = "CumPaidLoss_D")
}
```

### Mack Method

```{r wc_mack, results = "asis"}
wc_mack <- lapply(paid_tri[1:20], MackChainLadder, est.sigma = "Mack")
```

```{r}
mack_smry <- vector("list", length(wc_mack))
for (i in seq_along(wc_mack)) {
  mack_smry[[i]] <- summary(wc_mack[[i]])[[2]]$Totals[3:5]
}
```