---
title: "ChainLadder Package Projections in Action"
author: "Andy Merlino"
date: "2015-03-18"
output: html_document
fontsize: 12pt
geometry: margin=1.25in
---

[source code](https://raw.githubusercontent.com/merlinoa/ractuary-studies/master/chainladder-review.Rmd)

This analysis aims to measure the accuracy of the `ChainLadder::MackChainLadder()` and `ChainLadder::BootChainLadder()` reserve projection functions by comparing the projected reserves to actual reserves.

### Background

The [ChainLadder](http://cran.r-project.org/web/packages/ChainLadder/vignettes/ChainLadder.pdf) package by Markus Gesmann, Dan Murphy, and Wayne Zhang includes functions for estimating reserves.  `MackChainLadder()` provides a distribution free calculation of the standard error of the reserve projection.  `BootChainLadder()` uses a two stage bootstrap/simulation technique to determine the standard error of the reserve projection.

### Data

The CAS website provides historical U.S. insurance loss data across 6 different lines of business and many different companies.  Data from accident years 1988 through 1997 evaluated at each year end (i.e. the upper left triangle of losses) is used to run the `ChainLadder` functions.  In addition to the upper left triangle, the CAS data set includes actual losses for each accident year developed for 10 years; thus we can compare the projected reserve to the actual reserve after 10 development years.

For more information on the data see the [CAS webpage](http://www.casact.org/research/index.cfm?fa=loss_reserves_data).

```{r packages_data, message = FALSE, echo = FALSE}
library(ChainLadder) # install.packages("ChainLadder")
library(dplyr) # for data manipulation
library(reshape2) # data manipulation
library(ggplot2) # graphics
library(xtable) # tables
options(xtable.comment = FALSE)
options(xtable.type = "html")
library(shiny)

# package I created to store CAS data
library(casdata) # devtools::install_github("merlinoa/casdata")
# determine the minimum paid amount that we will consider
paid_cutoff <- 200
```

The following table breaks down the number of companies included in this analysis by line of business.

```{r data_cleaning, echo = FALSE, results = "asis"}
# Several companies were removed from the analysis because they 
# had incomplete or very few losses at one or more evaluation points.

# identify unique companies that have missing data.
# I also had some trouble with companies that had very small losses, so I decided
# to remove all companies with cumulative paid losses less than some value 
# at any observation in the data set.
# GRCODE is only unique by lob, so I need to create new column with combination of

df <- mutate(losses, id = paste(lob, gr_code, sep = "_"))
  
id_missing <- unique(df$id[df[, "paid"] <= paid_cutoff])

# remove companies that have missing data
df <- df[-which(df$id %in% id_missing), ]

df_out <- group_by(df, lob) %>%
  summarise(companies = length(unique(gr_code)))

names(df_out) <- c("Line of Business", "Number of Companies")
df_out[, 1] <- c("Commercial Auto", "Medical Malpractice",
                "Other Liabiltiy", "Private Passenger",
                "Product Liability", "Workers Compensation")

df_out <- xtable(as.data.frame(df_out))

print(df_out,
      include.rownames = FALSE)
```

### Analysis

I run the `MackChainLadder()` and `BootChainLadder()` reserve projection functions on each company.  I subtract the projected reserve from the actual reserve to arrive at the projection error.  To normalize the projection error, I divide the projection error by the estimated projection standard error.  Put a little more formally, I am simply interested in the normalized projection error, $Z$, where:

$Z = \frac{(p - \hat{p})}{\hat{se(\hat{p})}}$

Where: 

* $\hat{p}$ is the projected reserve.
* $\hat{se(\hat{p})}$ is the estimated standard error of the projected reserve. 
* $p$ is the actual reserve.  Where the actual reserve is the paid or incurred amount after 10   development years less the paid or incurred amount at the time of the projection estimate.

The following histograms should provide a visual interpretation of how the reserve projection errors are distributed around the reserve projection.

```{r create_triangles, echo = FALSE}
# Note: Our latest loss data is not necessarily developed to the ultimate loss
# amounts.  It is developed for 10 years; thus the paid and incurred loss values at
# the 10th development year may not be equal to one another.  Due to this possible
# discrepancy we define the actual reserve for projections based on paid data as the
# paid loss at the 10th development year less the paid loss at the time of the
# projection.  But for projections based on incurred data, we define the actual
# reserve as the incurred loss at the 10th development year less the incurred loss at
# the time of the projection.

# function: to create triangles for each company
# param data the cleaned insurance loss data frame; `df`
# param loss_type character "CumPaidLoss_D" or "IncurLoss_D"
triangles <- function(data, loss_type) {

 # find unique GRCODEs for looping
 unique_id <- unique(data$id)
 
 # create triangles for each company using `ChainLadder` package
 tri <- vector("list", length(unique_id))
 for (i in seq_along(unique_id)) {
   tri[[i]] <- as.triangle(data[data$id == unique_id[i] & 
                                      data$origin + 
                                      data$dev < 1999, ], 
                               origin = "origin", 
                               dev = "dev", 
                               value = loss_type)
 }
 # set name of each triangle to the id
 names(tri) <- unique_id
 tri
}

# use above functions to create paid and incurred triangles for all lines
# of business
paid_triangles <- triangles(df ,loss_type = "paid")
incurred_triangles <- triangles(df, loss_type = "incurred")
```

```{r projection_functions, echo = FALSE, warning = FALSE}
# function for mack and boot projections============================================
# param triangles list of all company triangles
# param method character "MackChainLadder" or "BootChainLadder"
projection <- function(triangles, method) {
  # run mack model as provided in ChainLadder package 
  projections <- lapply(triangles, method)
  
  # extract the good stuff from the boot projection
  smry <- vector("list", length(projections))
  for (i in seq_along(projections)) {
    smry[[i]] <- summary(projections[[i]])[[2]]
  }

  smry <- as.data.frame(t(as.data.frame(smry)))
  rownames(smry) <- NULL
  smry <- data.frame(names(triangles), smry)
  smry$lob <- gsub("_.*$", "", smry[, 1])
  smry
}
```

```{r projection_actual_join, echo = FALSE}
# function: to find the sum of actual losses for all origin years for 
# each insurance company after 10 development periods
# param data the cleaned insurance loss data frame; `df`
# param loss_type character "CumPaidLoss_D" or "IncurLoss_D"
actual_ultimate <- function(data, loss_type) {
  # find actual losses at development period 10
  dev_10 <- data[data$dev == 10, ]
  dev_10_sum <- as.data.frame(tapply(dev_10[, loss_type], dev_10$id,  sum))
  dev_10_sum$id <- rownames(dev_10_sum)
  names(dev_10_sum)[1] <- paste0(loss_type, "_ultimate")
  dev_10_sum
}

# run function to find actual ultimates for all companies
paid_actual_ultimate <- actual_ultimate(df, loss_type = "paid")
incurred_actual_ultimate <- actual_ultimate(df, loss_type = "incurred")

# group actual losses with mack projections
projection_actual_join <- function(projection_smry, actual) {
  smry <- left_join(projection_smry, actual, by = "id", copy = TRUE)
  names(smry)[length(smry)] <- "actual_ultimate"
  # calculate z value
  smry$actual_ibnr <-  smry$actual_ultimate - smry$latest
  smry$z = (smry$actual_ibnr - smry$ibnr) / smry$se
  smry
}
```

### Histograms

```{r mack_projections, echo = FALSE, warning = FALSE, message = FALSE}
# calculate mack projections and grab summary
paid_mack_smry <- projection(paid_triangles, 
                             method = "MackChainLadder")[, c(1, 2, 4, 5, 6, 8)]
names(paid_mack_smry) <- c("id", "latest", "ultimate", "ibnr", "se", "lob")

incurred_mack_smry <- projection(incurred_triangles, 
                                 method = "MackChainLadder")[, c(1, 2, 4, 5, 6, 8)]
names(incurred_mack_smry) <- c("id", "latest", "ultimate", "ibnr", "se", "lob")

# combine paid mack projections with actual
paid_mack_full <- projection_actual_join(paid_mack_smry, paid_actual_ultimate)

# combine incurred mack projections with actuals
incurred_mack_full <- projection_actual_join(incurred_mack_smry, 
                                      incurred_actual_ultimate)
```

```{r boot_projections, echo = FALSE, warning = FALSE, message = FALSE}
# caculate and return summary info for bootstrap method
paid_boot_smry <- projection(paid_triangles, method = "BootChainLadder")[, c(1, 2:5, 8)]
names(paid_boot_smry) <- c("id", "latest", "ultimate", "ibnr", "se", "lob")

incurred_boot_smry <- projection(incurred_triangles, method = "BootChainLadder")[, c(1, 2:5, 8)]
names(incurred_boot_smry) <- c("id", "latest", "ultimate", "ibnr", "se", "lob")

# combine paid mack projections with actual
paid_boot_full <- projection_actual_join(paid_boot_smry, paid_actual_ultimate)

# combine incurred boot projections with actual
incurred_boot_full <- projection_actual_join(incurred_boot_smry, 
                                      incurred_actual_ultimate)
```

```{r histogram_plots, echo = FALSE, message = FALSE, warning = FALSE}
# plot paid mack z values
ggplot(paid_mack_full, aes(x = z, fill = lob)) +
  geom_histogram(colour = "black") +
  xlab("Normalized Projection Error - Z") +
  ggtitle("Mack Projection - Paid Basis") +
  geom_bar()

# plot incurred mack z values
ggplot(incurred_mack_full, aes(x = z, fill = lob)) +
  geom_histogram(colour = "black") +
  xlab("Normalized Projection Error - Z") +
  ggtitle("Mack Projection - Incurred Basis") +
  geom_bar()

# plot paid boot z values
ggplot(paid_boot_full, aes(x = z, fill = lob)) +
  geom_histogram(colour = "black") +
  xlab("Normalized Projection Error - Z") +
  ggtitle("Boot Projection - Paid Basis") +
  geom_bar()

# plot incurred boot z values
ggplot(incurred_boot_full, aes(x = z, fill = lob)) +
  geom_histogram(colour = "black") +
  xlab("Normalized Projection Error - Z") +
  ggtitle("Boot Projection - Incurred Basis") +
  geom_bar()
```

### Initial Take Aways

Both the Mack and Boot paid reserve projections appear to slightly overestimate the actual reserve across all lines of business.  Several factors could be responsible for this overestimation (e.g. higher inflation in the years used to estimate the reserve than in the years after 1997).

The incurred normalized projection errors tend to be smaller in absolute value terms than the paid reserve projections.

The `BootChainLadder()` function on incurred losses gave significantly more accurate reserve projections than what would be expected if the projection errors were from the standard normal distribution.